---
title: "Homework6"
author: "Yunjia Liu"
date: "2024-12-02"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(broom)
```

## Probelm 1

1. Load the dataset.

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```


2. A function to calculate the r-squared and log(beta0 * beta1) \
First, establish a linear model between t_max and t_min. Then, we calculate the desired coefficient.

```{r}
compute_stats = function(data) {
  model = lm(tmax ~ tmin, data = data)
  glance_model = broom::glance(model)
  tidy_model = broom::tidy(model)
  
  r_squared = glance_model$r.squared
  
  beta0 = tidy_model$estimate[tidy_model$term == "(Intercept)"]
  beta1 = tidy_model$estimate[tidy_model$term == "tmin"]
  
  log_beta0_beta1 = log(abs(beta0 * beta1))
  return(tibble(r_squared = r_squared, log_beta0_beta1 = log_beta0_beta1))
}
```

3. A quick function to generate our bootstrap samples
```{r}
boot_sample = function(df) {
  sample_frac(df, replace = TRUE)
}
```

4. Perform bootstrap and compute r^2 and log(beta0 * beta1)
```{r}
n_boot = 5000
set.seed(123)

boot_straps =
  tibble(strap_number = 1:n_boot) |>
  mutate(
    strap_sample = map(strap_number, ~ boot_sample(weather_df)),  
    stats = map(strap_sample, compute_stats)
  )
```


5. Extract results into a tidy format.

```{r}
boot_results = 
  boot_straps |>
  unnest_wider(stats)
```

6. Calculate the 95% confidence intervals for r^2 and log(beta0 * beta1).

```{r}
ci_r_squared = quantile(boot_results$r_squared, c(0.025, 0.975))
ci_log_beta0_beta1 = quantile(boot_results$log_beta0_beta1, c(0.025, 0.975))

cat("95% CI for r^2: ", ci_r_squared, "\n")
cat("95% CI for log(beta0 * beta1): ", ci_log_beta0_beta1, "\n")
```

7. Plot distributions of two estimates (r^2 and log(beta0 * beta1))

```{r}
ggplot(boot_results, aes(x = r_squared)) +
  geom_histogram(binwidth = 0.01, color = "black", fill = "blue", alpha = 0.7) +
  labs(title = "Bootstrap Distribution of r^2", x = "r^2", y = "Frequency") +
  theme(
    plot.title = element_text(hjust = 0.5)
  )

ggplot(boot_results, aes(x = log_beta0_beta1)) +
  geom_histogram(binwidth = 0.05, color = "black", fill = "red", alpha = 0.7) +
  labs(title = "Bootstrap Distribution of log(beta0 * beta1)", x = "log(beta0 * beta1)", y = "Frequency") +
  theme(
    plot.title = element_text(hjust = 0.5)
  )
```

## Problem 2

```{r}


```


## Problem 3

```{r}

```

